---
title: "World Bank Assessment - STC Position"
author: Baqir Fateh
date: "`r format(Sys.time(), '%B %d, %Y')`"
format: html
editor: visual
---

```{r}
#| label: load-packages and download the raw data
#| meeage: false
#| warning: false
library(tidyverse)
library(Hmisc)
library(waldo)
library(matrixStats) # I used this package to calucate the weighted median 
library(DT)
library(rmarkdown)
#### Set up the Url to download the raw data 
tag      <- "202311081903"
base_url <- "https://github.com/randrescastaneda/pub_data/raw/"
data_url <- paste0(base_url, tag, "/data/Rtest1/")

## Download the raw data set for taks 1-3
wdi <- readr::read_rds(paste0(data_url, "wdi_in1.Rds"))
```

### 1. Summary statistics of GDP per capita by region

The replicated results in this task are almost identical to the original results.

```{r}
#| warning: false
############################################################
### 1. Summary statistics of GDP per capita by region      #
############################################################
q1_rep <- wdi%>%
  group_by(region, date)%>%
  summarise(N = n(), Mean = weighted.mean(gdp, pop, na.rm = TRUE),
            SD = sqrt(wtd.var(gdp, pop,na.rm = TRUE)),
            Min = min(gdp, na.rm = TRUE), Max = max(gdp, na.rm = TRUE))%>%
  rename(year = date)%>%
  as.data.frame()

## Display the output 
datatable(q1_rep, options = list(pageLength = 10), filter = "top")
```

### 2. Aggregate stats

The replicated results in this task are pretty close to the original outputs except for the median. I tried to estimate the weighted median using an existing function, but the estimates are off.

```{r}
#| warning: false
############################################
### 2. Aggregate stats                     #     
############################################
# List the relevant variables
vars <- c("lifeex", "gdp", "pov_intl")

# Initialize a list to store the results 
result <- list()

## Loop over the selected variables 
for (var in vars){
 q2_rep <- wdi%>%
    group_by(region, date)%>%
    summarise(mean = weighted.mean(!!sym(var), pop, na.rm = TRUE),
              sd = sqrt(wtd.var(!!sym(var), pop,na.rm = TRUE)),
              median = weightedMedian(!!sym(var), pop),
              pop = sum(pop, na.rm = TRUE),
              min = min(!!sym(var), na.rm = TRUE), 
              max = max(!!sym(var), na.rm = TRUE))%>%
    gather("estimate", !!var, -c(pop, region, date))
 result[[var]] <- q2_rep
}

## Combine the outputs to generate the final tables 
q2_rep <- Reduce(function(x, y) merge(x, y, by = c("region", "date", "estimate", "pop"), 
                                      all = TRUE), result)%>%
  as.data.frame()

## Display the output 
datatable(q2_rep, options = list(pageLength = 10), filter = "top")

## Download the answer table
#wdi_agg_out <- readr::read_rds(paste0(data_url, "wdi_agg_out.Rds"))%>%
  #as.data.frame()

```

### 3. Find outliers

The replicated results are pretty close to the original outputs.

```{r}
############################################
### 3. Find outliers                       #
############################################

q3_rep <- wdi%>%
  group_by(date)%>%
  mutate(mean_lifeex = weighted.mean(lifeex, pop, na.rm = TRUE),
         sd_lifeex = sqrt(wtd.var(lifeex, pop, na.rm = TRUE)),
         hl_lifeex = lifeex > mean_lifeex+2.5*sd_lifeex,
         ll_lifeex = lifeex < mean_lifeex-2.5*sd_lifeex)%>%
  mutate(mean_gdp = weighted.mean(gdp, pop, na.rm = TRUE),
         sd_gdp = sqrt(wtd.var(gdp, pop, na.rm = TRUE)),
         hl_gdp = gdp > mean_gdp+2.5*sd_gdp,
         ll_gdp = gdp < mean_gdp-2.5*sd_gdp)%>%
  mutate(mean_gini = weighted.mean(gini, pop, na.rm = TRUE),
         sd_gini = sqrt(wtd.var(gini, pop, na.rm = TRUE)),
         hl_gini = gini > mean_gini+2.5*sd_gini,
         ll_gini = gini < mean_gini-2.5*sd_gini)

## Display the output 
datatable(q3_rep, options = list(pageLength = 10), filter = "top")

## Visualize life-expectancy by year
q3_rep%>%
  group_by(date)%>%
  ggplot(aes(date,lifeex, mean_lifeex, col = region))+
  geom_point()+
  geom_line(aes(date, mean_lifeex), col = "blue", size = 0.7)+
  geom_ribbon(aes( ymin = mean_lifeex-2.5*sd_lifeex, ymax = mean_lifeex+2.5*sd_lifeex), 
              fill = "grey", color = NA, alpha = 0.6)+
  theme_minimal()+
  theme(legend.position = c(0.5, 0.16))+
  theme(legend.title = element_blank())+
  guides(color = guide_legend(ncol = 4, byrow = TRUE))
  
```

### 4. Poverty Measure

The replicated results are almost identical to the original outputs.

```{r}
#######################################
### 4. Poverty Measure                #
#######################################
## Read the raw data from github 
l_svy <- readr::read_rds(paste0(data_url, "svy_sim_in1.Rds"))

### Convert the list into a data frame 
## Add a year column
l_svy <- lapply(names(l_svy), function(year) {
  df <- l_svy[[year]]
  df$year <- year
  return(df)
})

## combine by row
sim_data <- do.call(rbind, l_svy)%>%
  mutate(year = str_sub(year,2,5))%>%
  as.data.frame()

## The setup of the output file 
# Poverty lines 
pov_lines <- c(2.15, 3.65, 6.85)

## Initialize a data frame to store the results
q4_rep <- expand.grid(year = unique(sim_data$year), pov_line = pov_lines)
q4_rep <- q4_rep%>%
  mutate(headcount = NA, povgap = NA, povseverity = NA)

## Write a customized function to compute the poverty indices 
pov_func <- function(income, pov_line, w, alpha){
  data <- data.frame(income, pov_line, w)
  data <- data%>%
    mutate(diff = ifelse(income < pov_line,((pov_line-income)/pov_line),0),
           pov_ind = ifelse(income < pov_line, diff^alpha,0))%>%
    summarise(pov_ind = sum(pov_ind*w),
              total_weight = sum(w))
  data$pov_ind/data$total_weight
}
## Compute poverty headcount, poverty gap, and poverty severity 
for (i in 1:nrow(q4_rep)){
  reduced_data <- sim_data%>%filter(year == q4_rep$year[i])
  q4_rep$headcount[i] <- pov_func(reduced_data$income, 
                                  pov_line = q4_rep$pov_line[i], 
                                  w = reduced_data$weight,alpha = 0)
  q4_rep$povgap[i] <- pov_func(reduced_data$income, 
                               pov_line = q4_rep$pov_line[i], 
                               w = reduced_data$weight,alpha = 1)
  q4_rep$povseverity[i] <- pov_func(reduced_data$income, 
                                    pov_line = q4_rep$pov_line[i], 
                                    w = reduced_data$weight,alpha = 2)
}

## Display the output 
datatable(q4_rep, options = list(pageLength = 10), filter = "top")

## Visualize the head count against year
q4_rep%>%
  ggplot(aes(year, headcount, group = pov_line, color = as.factor(pov_line)))+
  geom_line()+
  geom_point() +
  theme_minimal() +
  theme(legend.title = element_blank())+
  theme(legend.position = "bottom")

```

### 5. Lorenz curve

The replicated results in this task are different from the original output. However, the graph looks pretty similar.

```{r}
#######################################     
### 5. Lorenz curve                   #
#######################################
# Write a function 
lorenz_curve <- function(x){
  # Set up an empty data frame
  q5_rep <- data.frame(
    welfare = numeric(0),
    cum_welfare = numeric(0),
    cum_population = numeric(0),
    year = integer(0),
    bin = integer(0)
  )
  # Unique number of years 
  years <- unique(x$year)
  # Looping over each year
  for (i in seq_along(years)){
    yrs <- years[i]
    temp_out <- x %>%
      filter(year == yrs) %>%
      arrange(income) %>%
      mutate(cum_welfare_w = cumsum(income * weight),
             cum_weight = cumsum(weight)) %>%
      # Normalizing the cumulative share of income and population
      mutate(cum_welfare = cum_welfare_w / sum(income * weight),
             cum_population = cum_weight / sum(weight))%>%
      select(-c(income, area, weight))
    # Creating percentiles 
    percentiles <- seq(0, 1, length.out = 100)
    # Calculating the cumulative welfare for each percentile 
    cum_welfare_percent <- approx(temp_out$cum_population, temp_out$cum_welfare, percentiles)$y
    # Calculating the welfare for each percentile 
    welfare_percent <- approx(temp_out$cum_population, temp_out$cum_welfare_w, percentiles)$y
    # Creating a another temporary data frame to update the values of variables created in the temp_out 
    aux_out_table <- data.frame(
      welfare = welfare_percent,        
      cum_welfare = cum_welfare_percent,    
      cum_population = percentiles,     
      year = rep(yrs, length(percentiles)),    
      bin = seq(1, 100)                 
    )
    # Creating the final output table
    q5_rep <- rbind(q5_rep, aux_out_table)
  } 
  # Return the output table
  return(q5_rep)
}

q5_rep <- lorenz_curve(sim_data)

## Display the output 
datatable(q5_rep, options = list(pageLength = 10), filter = "top")

### Visualize the cumulative share of income against the cumulative share of population 
q5_rep%>%
  ggplot(aes(cum_population, cum_welfare, group = year, color = as.factor(year)))+
  geom_line()+
  theme_minimal() +
  theme(legend.title = element_blank())+
  theme(legend.position = c(0.1, 0.55))

```

### 6. Estimating Gini Coefficient

The replicated results are almost identical to the original outputs.

```{r}
##########################################
### 6. Estimating Gini Coefficient       #
##########################################

## Write function that estimate Gini coefficient for each year
gini_func <- function(x) {
  # Create an empty data frame to store the estimates
  q6_rep <- data.frame(year = integer(), gini = numeric())
  ## loop over the year 
  for (yr in unique(x$year)) {
    # Prepare the data for estimating the Gini coefficient
    gini_data <- x %>%
      filter(year == yr) %>%
      arrange(income) %>%
      mutate(cum_weight = cumsum(weight / sum(weight)),
             cum_income = cumsum(income * weight / sum(income * weight)))
    # Compute the Gini coefficient
    G <- 1 - sum(2 * diff(c(0, gini_data$cum_weight)) * c(0, head(gini_data$cum_income, -1)))
    # Store the yearly estimate of the Gini coefficient
    q6_rep <- rbind(q6_rep, data.frame(year = yr, gini = G))
  }
  return(q6_rep)
}

## Generate the Gini coefficient by year 
q6_rep <- gini_func(sim_data)

## Display the output 
datatable(q6_rep, options = list(pageLength = 10), filter = "top")

## Visualize the estimates of Gini Coefficient by year 
q6_rep%>%
  ggplot(aes(year, gini, group = FALSE))+
  geom_line()+
  geom_point()+
  theme_minimal() 

```
